---
title: "04-Sapwood_Parameters"
output: html_document
---

# Key Wood Traits {-}

## `[wood_density]:` Wood density {-}

- ED variable name : `rho`
- ED units: `g cm-3` Note! Leaf density is calculated in `kg cm-3` because that makes perfect sense ... 
- ED variable id: 1000000055
- FATES varaiable name: `WD`

In the database, wood_density is unitless. 
So I'm assuming it can be directly mapped to our `wood_density` as such:

$$\frac{WD (gcm^{-2})}{\rho_w(gcm^{-2})} * \rho_w(gcm^{-2}) =  \frac{WD (gcm^{-2})}{1(gcm^{-2})} * 1(gcm^{-2})$$
There are multiple priors available for wood density so we are explicitly choosing prior `1000000281` (this was chosen by someone else when the PFT was created.)

```{r wood_density}
t = "wood_density"

traits[[t]] <- list(
  variable_name = t,
  variable_id =  pfts_priors %>% filter(var_name == t) %>% pull(variable_id),
  prior_id = pfts_priors %>% filter(var_name == t) %>% pull(id),
  fit = pfts_priors %>% filter(var_name == t) %>% select("distn", "parama", "paramb")
)

traits[[t]]$prior = rdistn(traits[[t]]$fit)
traits[[t]]$default = get_ED_default(PFT3_defaults_history, "rho")

prior_plot(prior = traits[[t]]$prior, 
           plot_default = traits[[t]]$default,
           title = sprintf("%s:\n(%s, %.4f, %.4f)", t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "prior")
```

--------------------------------------------------------------------------------

## `[wood_psi_osmotic]:` Sapwood osmotic water potential at saturation  {-}

### ED  {-}

- ED variable name : `wood_psi_osmotic`
- ED variable units: `m`

### BETY  {-}

- Variable id: `1000000298`
- Prior id: `1000000410`

NOTE!! The units are NEGATIVE in the database 
so we can calculate a distribution that is entirely negative. 

### Calculation  {-}

#### Expert elicitation {-}

Expert elicitation provided by Brad Christoffersen 

- FATES varaiable name: `pinot_node (stem)`
- FATES variable units: `m`

#### Conversion {-}
`MPa to m`

#### Code {-}


```{r calc wood_psi_osmotic}

t = "wood_psi_osmotic"
t_long = "Sapwood osmotic water potential at saturation "

i <- which(priors$ED_name == t)
prior_in <- priors[i, stats] * C$MPa2m

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)
```




```{r plot wood_psi_osmotic}

traits[[t]]$prior = -1 * rdistn(traits[[t]]$fit)
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "elic")

```

--------------------------------------------------------------------------------

## `[wood_elastic_mod]:` Sapwood bulk elastic modulus {-}

### ED  {-}

- ED variable name : `wood_elastic_mod`
- ED variable units: `MPa` Note! This is the only case in which MPa is used instead of m!

### BETY  {-}

- Variable id: `1000000303`
- Prior id: `1000000394`

### Calculation  {-}

#### Expert elicitation {-}

Expert elicitation provided by Brad Christoffersen 

- FATES varaiable name: `epsil_node (stem)` 
- FATES variable units: `MPa`

#### Conversion {-}

None!

### Code {-}

```{r calc wood_elastic_mod}

t = "wood_elastic_mod"
t_long = "wood bulk elastic modulus "

i <- which(priors$ED_name == t)
prior_in <- priors[i,]

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)

```


```{r plot wood_elastic_mod}

traits[[t]]$prior = rdistn(traits[[t]]$fit)
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
           q = c(0,.999), 
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "elic")

```

--------------------------------------------------------------------------------

## `[wood_water_sat]:` Sapwood water content at saturation {-}

### ED {-}

- ED variable name: `wood_water_sat`
- ED variable units: `kg H2O/kg biomass`

### BETY  {-}

- Variable id: ``
- Prior id: ``

### Calculation  {-}

#### Expert elicitation {-}

Expert elicitation provided by Brad Christoffersen 

- FATES variable name: `thetas_node (stem)`
- FATES variable units: `kg kg-1`

#### Conversion  {-}

multiply by the density of water over wood density

### Code {-}

```{r calc wood_water_sat}

t = "wood_water_sat"
t_long = "Wood water content at saturation "

i <- which(priors$ED_name == t)
temp_wws_fit <- prior_get_fit(priors[i,], accepted_dists, plot = FALSE)  

temp_wws_fit$score$dist[which.min(temp_wws_fit$score$RMSE)]
# I'm skipping steps here for the sake of time and sanity
# Normal is the best fit

temp_wws_prior <- rnorm(100000, temp_wws_fit$dists$norm[1], temp_wws_fit$dists$norm[2])

temp_wws_sample <- temp_wws_prior * (C$wdns/(traits[["wood_density"]]$prior * 1e3)) 
# Note that here we have to multiply by 1e3 because of the units of wood density 

# In this case I'm not using the function I built because here it wouldn't make sense. 
prior_in <- data.frame(
  theor.min = NA,
  low.025 =  quantile(temp_wws_sample, c(.025), na.rm = TRUE),
  low.25  =  quantile(temp_wws_sample, c(.25), na.rm = TRUE),
  mean    =  quantile(temp_wws_sample, c(.5), na.rm = TRUE), 
  upp.75  =  quantile(temp_wws_sample, c(.75), na.rm = TRUE), 
  upp.975 =  quantile(temp_wws_sample, c(.975), na.rm = TRUE),
  theor.max = NA)

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)

# Now we can remove temp variables
rm(temp_wws_fit, temp_wws_prior, temp_wws_sample)
```
```{r plot wood_water_sat}

traits[[t]]$prior = rdistn(traits[[t]]$fit)
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
           q = c(0,.99), 
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "mixed")

```

--------------------------------------------------------------------------------

## `[wood_psi_tlp]:` Sapwood water potential at turgor loss point {-}

### ED {-}

- ED variable name:  `wood_psi_tlp`
- ED variable units: `m`

### BETY  {-}

- Variable id: ``
- Prior id: ``

NOTE!! The units are NEGATIVE in the database 
so we can calculate a distribution that is entirely negative. 

### Calculation  {-}

#### Expert elicitation {-}

No expert elicitation provided by Brad Christoffersen. 

There is a FATES parameter but it is calculated from other parameters

- FATES variable name: `pitlp_node`
- FATES variable units: `MPa`

Constraints we added
- `wood_psi_tlp > .1 MPa` (this was given as a boundary for other psi variables)

#### Conversion  {-}

- Calculated from other parameters: `wood_psi_osmotic` and `wood_elastic_mod`

- Remember that `wood_elastic_mod` is `MPa`, where everything else is `m`! 

### Code {-}

```
wood_psi_tlp_samp = 
(wood_psi_osmotic_prior * wood_elastic_mod_prior * C$MPa2m)/
(wood_psi_osmotic_prior + wood_elastic_mod_prior * C$MPa2m)
```

```{r calc wood_psi_tlp_samp}

t = "wood_psi_tlp"
t_long = "Sapwood water potential at turgor loss point "


temp_wpt_sample <- 
  (traits[["wood_psi_osmotic"]]$prior * traits[["wood_elastic_mod"]]$prior * C$MPa2m)/
  (traits[["wood_psi_osmotic"]]$prior + traits[["wood_elastic_mod"]]$prior * C$MPa2m)

summary(temp_wpt_sample)

i <- which(priors$ED_name == t)
prior_in <- priors[i,stats] * C$MPa2m # Convert from MPa to m
prior_in <- prior_add_samp_data(prior_in, samp = -temp_wpt_sample)
prior_in[,stats]

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)
```

```{r plot wood_psi_tlp}

traits[[t]]$prior = -1 * rdistn(traits[[t]]$fit) # Remember to flip back over for psi
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
           q = c(.005,1),
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "mixed")

```

--------------------------------------------------------------------------------

## `[wood_water_cap]:` Wood hydaulic capacitance {-}

### ED  {-}

- ED variable name : `wood_water_cap`
- ED variable units: `kg H2O/kg biomass/m`

### BETY {-}

Different units in BETY so we can fit a distribution!
The numbers were so small that I couldn't get the fitting function to work until
I scaled the variable up!

- BETY varaible units: `g H2O / kg biomass / m`

### Calculation  {-}

#### Expert elicitation {-}

No expert elicitation provided by Brad Christoffersen. 

Constraints we added
- `wood_water_cap > 1e-07 kg H2O/kg biomass/m` 

### Code {-}

```{r calc wood_water_cap}
# wood_water_cap_samp <- 
#   1000 * (1/wood_density_prior) * (1 - wood_psi_osmotic_prior / (4 * wood_psi_tlp_prior)) * 
#   (wood_water_sat_prior / (4 * abs(wood_psi_tlp_prior))) 

temp_wwc_samp <- 
  1000 * (1/traits[["wood_density"]]$prior) * (1 - traits[["wood_psi_osmotic"]]$prior / 
                                                 (4 * traits[["wood_psi_tlp"]]$prior)) * 
  (traits[["wood_water_sat"]]$prior / 
     (4 * abs(traits[["wood_psi_tlp"]]$prior))) 

summary(temp_wwc_samp)

t = "wood_water_cap"
t_long = "Wood Water Cap "

i <- which(priors$ED_name == t)
prior_in <- priors[i,stats] 
prior_in <- prior_add_samp_data(prior_in, samp = temp_wwc_samp)
prior_in[,stats]

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)
```
```{r plot wood_water_cap}

# I don't actually remember why we were plotting 10, 100, 300, 900 ... ?

traits[[t]]$prior = rdistn(traits[[t]]$fit) * (1/1000) # Convert back to kg
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
           q = c(0, .975),
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "mixed")  +
  geom_vline(aes(xintercept = 10/C$MPa2m * (1/1000)), color = "gray") +
  geom_vline(aes(xintercept = 100/C$MPa2m * (1/1000)), color = "gray") +
  geom_vline(aes(xintercept = 300/C$MPa2m * (1/1000)), color = "gray") +
  geom_vline(aes(xintercept = 900/C$MPa2m * (1/1000)), color = "gray")

```

--------------------------------------------------------------------------------

## `[wood_psi_min]:` Sapwood minimum relative water content or Sapwood residual fraction {-}

### ED {-}

- ED variable name : `wood_psi_min`
- Ed variable units: `m`

### BETY {-}

Negative units in BETY so we can fit a distribution! 

- BETY varaible units: `-m`

### Calculation  {-}

#### Expert elicitation {-}

There are no FATES parameters 

Constraints we added
- `wood_psi_min > .1 MPa` (this was given as a boundary for other psi variables)

### Code {-}

Calculated from other priors. 

Should I put `wood_rwc_min` and `rwc_tlp_wood` in the database too?

```{r}
# rwc_tlp_wood_samp <- 1. - (1. - 0.75 * wood_density_prior) / (2.74 + 2.01 * wood_density_prior) 
# wood_rwc_min_samp = wood_elastic_mod_prior * (1. - rwc_tlp_wood_samp) / (wood_psi_osmotic_prior / C$MPa2m) + 1.
# wood_psi_min_samp = (wood_rwc_min_samp - 1.) * wood_water_sat_prior / wood_water_cap_prior

temp_rwc_tlp_wood_samp <- 1. - (1. - 0.75 * traits[["wood_density"]]$prior) / (2.74 + 2.01 * traits[["wood_density"]]$prior)
temp_wood_rwc_min_samp = traits[["wood_elastic_mod"]]$prior * (1. - temp_rwc_tlp_wood_samp) / (traits[["wood_psi_osmotic"]]$prior / C$MPa2m) + 1.
temp_wood_rwc_min_samp = (temp_wood_rwc_min_samp - 1.) * traits[["wood_water_sat"]]$prior / traits[["wood_water_cap"]]$prior


t = "wood_psi_min"
t_long = "Wood Psi Min "

i <- which(priors$ED_name == t)
prior_in <- priors[i,stats] * C$MPa2m # Convert from MPa to m
prior_in <- prior_add_samp_data(prior_in, samp = -temp_wood_rwc_min_samp)
prior_in[,stats]

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)


```

```{r}
traits[[t]]$prior = -1 * rdistn(traits[[t]]$fit) # Remember to flip back over for psi
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
          q = c(.025,1),
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "mixed")
```

--------------------------------------------------------------------------------

## `[wood_Kmax]:` Maximum hydraulic conductivity of the stem {-}

### ED {-}

- ED variable name : `wood_Kmax`
- ED variable units: `kg H2O / m / s` Note! This is the only case in which MPa is used instead of m!

### BETY {-}

### Calculation  {-}

#### Expert elicitation {-}

Expert elicitation provided by Brad Christoffersen 

- FATES varaiable name: `kmax_node` 
- FATES variable units: `kg m-1 s-1 MPa-1` 

#### Conversion {-}

I don't think we need to multiply by the density of water  ... which is confusing. 

### Code {-}

```{r calc wood_Kmax}
t = "wood_Kmax"
t_long = "Maximum hydraulic conductivity of the stem "

i <- which(priors$ED_name == t)
prior_in <- priors[i, stats] / C$MPa2m 

fit_trait_dist(bety, traits, t, prior_in, accepted_dists, return_traits = FALSE)

# In this case, the exponential function is the best fit, 
# but I'm going to choose the beta function because it is much easier to deal 
# with in the meta analysis. This could be wrong and we may need to return to it. 
traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists = "beta")

```

```{r plot wood_Kmax}

traits[[t]]$prior = rdistn(traits[[t]]$fit) 
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
          q = c(0,.975),
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "elic")

```

--------------------------------------------------------------------------------

## `[wood_psi50]:` Water potential at which 50% of stem conductivity is lost {-}

### ED {-}

- ED variable name : `wood_psi50`
- ED variable units: `m` 

### BETY {-}

Negative units in BETY so we can fit a distribution! 

- BETY varaible units: `-m`

### Calculation {-}

NOTE: There is an *ERROR* in the ED code for the equation for `wood_psi50`

Should be:

`(-1.09-(3.57 * rho(ipft) ^ 1.73) * C$MPa2m`

This one is tricky, do we keep the normal distribution because it's the best fit?

#### Expert elicitation {-}

- FATES varaiable name: `p50_node (stem)`
- FATES variable units: `MPa`

#### Conversion {-}
`MPa to m`

### Code {-}

#### Using the best fit norm {-}
```{r calc wood_psi50}
t = "wood_psi50"
t_long = "Water potential at which 50% of stem conductivity is lost "

i <- which(priors$ED_name == t)
prior_in <- priors[i, stats] * C$MPa2m

fit_trait_dist(bety, traits, t, prior_in, accepted_dists, return_traits = FALSE)

# In this case norm is the best fitting ditribution, 
# But a positive number ends up falling within the 95% CI of the distribution.
# So I'm going to choose a log normal distribution even though it is not the best fit. 
# Hopefully this can also be fixed given that we have data for both psi50 and Kexp... 

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists = "lnorm")

```

```{r plot wood_psi50}
traits[[t]]$prior = -1 * rdistn(traits[[t]]$fit) 

# Calculating default from wood density by hand because it is calculated incorrectly in ED! 
traits[[t]]$default = (-1.09 - (3.57 * traits[["wood_density"]]$default ^ 1.73)) * C$MPa2m

prior_plot(prior = traits[[t]]$prior, 
          q = c(0.025,1),
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "elic")
```


--------------------------------------------------------------------------------

## `[ax]:` Slope of xylem vulnerability curve at P50 {-}

### ED {-}

Not in ED, but will be used to calculate `Kexp`.

### BETY {-}

### Calculation {-}

#### Expert elicitation {-}

Expert elicitation provided by Brad Christoffersen 

- FATES varaiable name: `avuln_node (stem)` 
- FATES variable units: unitless

#### Conversion {-}

None! 

### Code {-} 

```{r calc ax}

t = "ax"
t_long = "Slope of xylem vulnerability curve at P50 "
i <- which(priors$Christoffersen_name == "avuln_node")
prior_in <- priors[i, stats]

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)

```

```{r plot ax}
traits[[t]]$prior = rdistn(traits[[t]]$fit) 

prior_plot(prior = traits[[t]]$prior, 
          q = c(0,.975),
           title = sprintf("%s(%s):\n(%s, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama),
          type = "elic")

```


## `[wood_Kexp]:` Exponent for the hydraulic vulnerability curve of stem conductivity  {-}
... under the Weibull function `1/(1+(psi/psi50) ^ Kexp_stem)`

### ED {-}

- ED variable name : `wood_Kexp`
- ED variable units: unitless 

### BETY {-}

### Calculation {-}

Kexp = -ax * 4 * p50 / 100

#### Expert elicitation {-}

None

#### Conversion {-}

None! 

### Code {-} 

```{r calc Kexp}

t = "wood_Kexp"
t_long = "Exponent for the hydraulic vulnerability curve of stem conductivity  "

temp_wood_Kexp_samp <- -1 * traits[["ax"]]$prior * 4 * (traits[["wood_psi50"]]$prior/C$MPa2m) / 100 
summary(temp_wood_Kexp_samp)

i <- which(priors$ED_name == t)
prior_in <- priors[i,stats] 
prior_in <- prior_add_samp_data(prior_in, samp = temp_wood_Kexp_samp)
prior_in[,stats]

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)

```

```{r plot Kexp}

traits[[t]]$prior = rdistn(traits[[t]]$fit) 

# Calculating default from wood density by hand because it is calculated incorrectly in ED! 
traits[[t]]$default = 0.544 * 4. * (-traits[["wood_psi50"]]$default / C$MPa2m) ^ (-0.17)

prior_plot(prior = traits[[t]]$prior, 
                q = c(0,.975), 
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "equation")
```

Note: here the miscalculation of psi50 in the code barely shows up 
because the value is divided by a large number. 
