---
title: "07-Hydraulics_Leaf"
output: html_document
---

# Hydraulic Leaf Traits {-}

## `[leaf_psi_osmotic]:` Leaf osmotic water potential at saturation {-}

### ED  {-}

- ED variable name : `leaf_psi_osmotic`
- ED variable units: `m`

### BETY  {-}

NOTE!! The units are NEGATIVE in the database 
so we can calculate a distribution that is entirely negative. 

- Variable id: `1000000295` 
- Prior id: `1000000388`

### Calculation  {-}

#### Expert elicitation  {-}

Expert elicitation provided by Brad Christoffersen 

- FATES varaiable name: `pinot_node (leaf)`
- FATES variable units: `MPa`

#### Conversion  {-}
`MPa to m`

### Code {-}

```{r calc leaf_psi_osmotic}

t = "leaf_psi_osmotic"
t_long = "Leaf osmotic water potential at saturation "

i <- which(priors$ED_name == t)
prior_in <- priors[i, stats] * C$MPa2m

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)
```

```{r plot leaf_psi_osmotic}

traits[[t]]$prior = -1 * rdistn(traits[[t]]$fit)
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "elic")
```

--------------------------------------------------------------------------------

## `[leaf_elastic_mod]`: Leaf bulk elastic modulus {-}

### ED  {-}

- ED variable name : `leaf_elastic_mod`
- ED variable units: `MPa` Note! This is the only case in which MPa is used instead of m!

### BETY  {-}

- Variable id: `1000000294`
- Prior id: `1000000387`

### Calculation  {-}

#### Expert elicitation {-}

Expert elicitation provided by Brad Christoffersen 

- FATES varaiable name: `epsil_node (leaf)` 
- FATES variable units: `MPa`

#### Conversion {-}

None!

### Code {-}

```{r calc leaf_elastic_mod}

t = "leaf_elastic_mod"
t_long = "Leaf bulk elastic modulus "

i <- which(priors$ED_name == t)
prior_in <- priors[i,]
prior_in

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)

```


```{r plot leaf_elastic_mod}

traits[[t]]$prior = rdistn(traits[[t]]$fit,1e06)
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
           # q = c(0,.999), 
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "elic")
```

--------------------------------------------------------------------------------

## `[leaf_density]:` Density of leaf tissue {-}

### ED  {-}

- ED variable name : `leaf_density`
- ED variable units: `kg m-3`

### BETY  {-}

- Variable id: `1000000304`
- Prior id: `1000000396`

### Calculation  {-}

#### Expert elicitation {-}

No expert elicitation provided by Brad Christoffersen. 

I will use the equation in ED to calcualte the prior and 
add min and max values to make sure the distribution 
does not predict values that are not biologically possible. 

Copies directly from the ED code: `max(0.1 * 1.e3, (leaf_elastic_mod(ipft) - 2.03) / 25.4 * 1.e3)`

But there is a key step here that we need to think about, which is that in the ED code, `0.1 * 1.e3` is the smallest possible value.

Trying to set `0.1 * 1.e3` as the min theoretical value when fitting the distribution (but remove the `max` from the equation when calculating the sample prior.) resulted in a poorly fitting distribution. 

Thus I'm going to set 1e-07 as the theoretical minimum value, fit a distribution and then exclude values smaller than `0.1 * 1.e3` after sampling from the distribution.  

`leaf_density_sample <- (leaf_elastic_mod_prior - 2.03) / 25.4 * 1.e3`

Values added by Betsy and Felicien: 

```
# Leaf_density
priors[which(priors$ED_name == "leaf_density"),
       c("theor.min", "theor.max")] <- c(1e-7, 2000)

```

#### Conversion {-}

None

### Code {-}

```{r calc leaf_density_sample}

t = "leaf_density"
t_long = "Density of leaf tissue "

i <- which(priors$ED_name == "leaf_density")

temp_leaf_density_sample <- (traits[["leaf_elastic_mod"]]$prior - 2.03) / 25.4 * 1.e3

# Add in the max an min from the priors table
prior_in <- priors[i,] # Note here I don't need to do any units conversion 
prior_in <- prior_add_samp_data(prior_in, samp = temp_leaf_density_sample)
prior_in[,stats]
# # I think I fucked up the theoretical mean ... it should have been
# prior_in[,"theor.min"] <- 0.1 * 1.e3
prior_in[,"theor.min"] <- 1e-7
prior_in[,"theor.max"] <- NA

# prior_in[,stats]
# prior_in[,stats]

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)

```

```{r plot leaf_density_sample}

traits[[t]]$prior = rdistn(traits[[t]]$fit, 1000000)
# Here I actually I think I will 
traits[[t]]$prior = sample(x = traits[[t]]$prior[which(traits[[t]]$prior > 100)], size = 1e5)

traits[[t]]$default = (get_ED_default("/fs/data3/ecowdery/ED.Hydro/parameters/pft3_defaults_history.xml", "leaf_elastic_mod") - 2.03) / 25.4 * 1.e3

prior_plot(prior = traits[[t]]$prior, 
           # q = c(0,.995), 
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "mixed")
```

--------------------------------------------------------------------------------

## `[leaf_water_sat]:` Leaf water content at saturation {-}

### ED  {-}

- ED variable name : `leaf_water_sat`
- ED variable units: `kg H2O/kg biomass`

### BETY  {-}

- Variable id: `1000000285`
- Prior id: `1000000414`

### Calculation  {-}

There are a couple ways to approach this:

Either build a prior completely from other priors 
(ie using the equations from ED and a prior for LMA and leaf_density) 
or try to do a combination where we use the priovided expert elicitation 
and the prior distribution for leaf_density. 

I am going to try doing both just to see how they differ. 

#### Expert elicitation {-}

Expert elicitation provided by Brad Christoffersen 

- FATES varaiable name: `thetas_node (leaf)`
- FATES variable units: `kg kg-1`

#### Conversion {-}

The equations between the two are identical except that to get from 
FATES to ED, must multiply by `(water density)/(leaf density)`

### Code {-}

#### Initial testing code {-}

```{r test leaf_water_sat, eval = FALSE, echo=FALSE, warning=FALSE}
# source("Leaf_water_sat_prior_or_elic.R")
```

From this point, we can see that just using the equations in the ED code makes 
the distributions predict low values which is could be a problem given that 
these values should not be zero and that the distribution underpredicts both 
in relation to the default value and the values that Felicien gave me. 

Thus I would create the prior using the expert elicitation values provided by Brad. 

#### Actual code {-}

```{r calc leaf_water_sat}
t = "leaf_water_sat"
t_long = "Leaf water content at saturation "
i <- which(priors$ED_name == t)
temp_myfit_0 <- prior_get_fit(priors[i,], accepted_dists, plot = FALSE)  

temp_myfit_0$score$dist[which.min(temp_myfit_0$score$RMSE)]
# I'm skipping steps here just for the sake of time & sanity
temp_leaf_water_sat_prior_0 <- rnorm(100000, temp_myfit_0$dists$norm[1], temp_myfit_0$dists$norm[2])

temp_leaf_water_sat_sample <- temp_leaf_water_sat_prior_0 * (C$wdns/(traits[["leaf_density"]]$prior))

# plot(density(temp_leaf_water_sat_sample), main = "leaf water sat sample without fitting")

# In this case I'm not using the function I built because here it wouldn't make sense. 
prior_in <- data.frame(
  theor.min = NA,
  low.025 =  quantile(temp_leaf_water_sat_sample, c(.025), na.rm = TRUE),
  low.25  =  quantile(temp_leaf_water_sat_sample, c(.25), na.rm = TRUE),
  mean    =  quantile(temp_leaf_water_sat_sample, c(.5), na.rm = TRUE), 
  upp.75  =  quantile(temp_leaf_water_sat_sample, c(.75), na.rm = TRUE), 
  upp.975 =  quantile(temp_leaf_water_sat_sample, c(.975), na.rm = TRUE),
  theor.max = NA)

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)
```


```{r plot leaf_water_sat}

traits[[t]]$prior = rdistn(traits[[t]]$fit)
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
           q = c(0,.995), 
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "mixed")

```


--------------------------------------------------------------------------------

## `[leaf_psi_tlp]:` Leaf water potential at turgor loss point {-}

### ED  {-}

- ED variable name : `leaf_psi_tlp`
- ED variable units: `m`

### BETY  {-}

- Variable id: `1000000284`
- Prior id: `1000000384`

NOTE!! The units are NEGATIVE in the database 
so we can calculate a distribution that is entirely negative. 

### Calculation  {-}

#### Expert elicitation {-}

No expert elicitation provided by Brad Christoffersen. 

I will use the equation in ED to calcualte the prior and 
add min and max values to make sure the distribution 
does not predict values that are not biologically possible. 

Values added by Betsy and Felicien: 

```
# Leaf_psi_tlp
priors[which(priors$ED_name == "leaf_psi_tlp"),
      c("theor.min", "theor.max")] <- c(-6, -0.5)
```

Code directly from ED: `leaf_elastic_mod(ipft)  * (leaf_psi_osmotic(ipft) / C$MPa2m) / (leaf_elastic_mod(ipft) + leaf_psi_osmotic(ipft) / C$MPa2m) * C$MPa2m`

Thus, `leaf_psi_tlp` is calculated purely from other parameters: `leaf_psi_osmotic` and `leaf_elastic_mod`

```
temp_leaf_psi_tlp_samp <- 
  (leaf_psi_osmotic_prior * (leaf_elastic_mod_prior * C$MPa2m))/
  (leaf_psi_osmotic_prior + (leaf_elastic_mod_prior * C$MPa2m))
```

#### Conversion {-}

Remember that `leaf_elastic_mod` is `MPa`, where everything else is `m`! 

### Code {-}

```{r calc leaf_psi_tlp}
t = "leaf_psi_tlp"
t_long = "Leaf water potential at turgor loss point "

temp_leaf_psi_tlp_samp <-
(traits[["leaf_psi_osmotic"]]$prior * (traits[["leaf_elastic_mod"]]$prior * C$MPa2m))/
(traits[["leaf_psi_osmotic"]]$prior + (traits[["leaf_elastic_mod"]]$prior * C$MPa2m))

# Add in the max an min from the priors table
i <- which(priors$ED_name == t)
prior_in <- priors[i,stats] * C$MPa2m # Convert from MPa to m
prior_in <- prior_add_samp_data(prior_in, samp = -temp_leaf_psi_tlp_samp)
prior_in[,stats]

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)
```


```{r plot leaf_psi_tlp}

traits[[t]]$prior = -1 * rdistn(traits[[t]]$fit) # Remember to flip back over for psi
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
           q = c(.005,1),
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "mixed")

# prior_plot(prior = traits[[t]]$prior * (-1/C$MPa2m), 
#            q = c(0,.995),
#            plot_default = traits[[t]]$default * (-1/C$MPa2m),
#            title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
#                            traits[[t]]$fit$distn, 
#                            traits[[t]]$fit$parama,
#                            traits[[t]]$fit$paramb),
#            type = "mixed")

```

--------------------------------------------------------------------------------

## `[leaf_water_cap]:` Leaf hydaulic capacitance. {-}

### ED  {-}

- ED variable name : `leaf_water_cap`
- ED variable units: `kg H2O/kg biomass/m`

### BETY  {-}

- Variable id: `1000000287`
- Prior id: `1000000405`

NOTE: units in BETY are `g H2O / kg biomass / m`! This is because we needed to multiply by 1000 to get large enough values to fit the distribution.

### Calculation  {-}

#### Expert elicitation {-}

No expert elicitation provided by Brad Christoffersen. 

I will use the equation in ED to calcualte the prior and 
add min and max values to make sure the distribution 
does not predict values that are not biologically possible. 

Values added by Betsy and Felicien: 

```
priors[which(priors$ED_name == "leaf_water_cap"),
       c("theor.min", "theor.max")] <- c(1e-7, NA)
```

Equation copied directly from ED: `(1. - leaf_psi_osmotic(ipft) / (4. * leaf_psi_tlp(ipft))) * leaf_water_sat(ipft) / (4. * abs(leaf_psi_tlp(ipft)))`

#### Conversion {-}

Don't forget about the switch from kg to g!

### Code {-}

```{r calc leaf_water_cap}
t = "leaf_water_cap"
t_long = "Leaf hydaulic capacitance "

# leaf_water_cap_samp <- 
#   1000 * (1 - leaf_psi_osmotic_prior / (4 * leaf_psi_tlp_prior)) * 
#   (leaf_water_sat_prior / (4 * abs(leaf_psi_tlp_prior))) # * (C$wdns/leaf_density_prior)
# 
# leaf_water_cap_samp <- leaf_water_cap_samp[0 < leaf_water_cap_samp]


temp_leaf_water_cap_samp <- 
  1000 * (1 - traits[["leaf_psi_osmotic"]]$prior / (4 * traits[["leaf_psi_tlp"]]$prior)) * 
  (traits[["leaf_water_sat"]]$prior / (4 * abs(traits[["leaf_psi_tlp"]]$prior))) # * (C$wdns/traits[["leaf_density"]]$prior)

temp_leaf_water_cap_samp <- temp_leaf_water_cap_samp[0 < temp_leaf_water_cap_samp]

# Add in the max an min from the priors table
i <- which(priors$ED_name == "leaf_water_cap")
prior_in <- priors[i,stats] 
prior_in[,"theor.min"] <- NA
# prior_in[,"theor.max"] <- max(.35 * (1/C$MPa2m) * (C$wdns/leaf_density_prior) )
prior_in <- prior_add_samp_data(prior_in, samp = temp_leaf_water_cap_samp)
prior_in[,stats]

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)

```


```{r plot leaf_water_cap}
traits[[t]]$prior = rdistn(traits[[t]]$fit) * (1/1000) # Convert back to kg
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
           q = c(0, .975),
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "mixed")  
```

--------------------------------------------------------------------------------

## `[leaf_psi_min]:` Leaf minimum water potential  {-}

This is calculated from `[leaf_rwc_min]:` Leaf minimum relative water content or leaf residual fraction, which I have not added to BETY. 

### ED  {-}

- ED variable name : `leaf_psi_min`
- ED variable units: `m`

### BETY  {-}

- Variable id: `1000000299`
- Prior id: `1000000392`

NOTE!! The units are NEGATIVE in the database 
so we can calculate a distribution that is entirely negative. 

### Calculation  {-}

#### Expert elicitation {-}

No expert elicitation provided by Brad Christoffersen. 

I will use the equation in ED to calcualte the prior and 
add min and max values to make sure the distribution 
does not predict values that are not biologically possible. 

Values added by Betsy and Felicien: 

```
# Leaf_psi_min
priors[which(priors$ED_name == "leaf_psi_min"),
       c("theor.min", "theor.max")] <- c(-700, -0.1) 
```

Code copied from ED: 

```
leaf_rwc_min = 0.01 * leaf_elastic_mod(ipft) + 0.17
leaf_psi_min = (leaf_rwc_min(ipft) - 1.) * leaf_water_sat(ipft) / leaf_water_cap(ipft)
```

#### Conversion {-}

None

### Code {-}

```{r calc leaf_psi_min}
t = "leaf_psi_min"
t_long = "Leaf minimum water potential "

# traits[["leaf_water_sat"]]$prior

temp_leaf_rwc_min_samp = 0.01 * traits[["leaf_elastic_mod"]]$prior + 0.17
temp_leaf_psi_min_samp = (temp_leaf_rwc_min_samp - 1.) * traits[["leaf_water_sat"]]$prior / traits[["leaf_water_cap"]]$prior

# Add in the max an min from the priors table
i <- which(priors$ED_name == "leaf_psi_min")
prior_in <- priors[i,stats] * C$MPa2m # Convert from MPa to m
prior_in <- prior_add_samp_data(prior_in, samp = -temp_leaf_psi_min_samp)
prior_in[,stats] / C$MPa2m

traits <- fit_trait_dist(bety, traits, t, prior_in, accepted_dists)

```

```{r plot leaf_psi_min}

traits[[t]]$prior = -1 * rdistn(traits[[t]]$fit) # Remember to flip back over for psi
traits[[t]]$default = get_ED_default(PFT3_defaults_history, t)

prior_plot(prior = traits[[t]]$prior, 
          q = c(.025,1),
           plot_default = traits[[t]]$default,
           title = sprintf("%s(%s):\n%s(%.4f, %.4f)", t_long, t, 
                           traits[[t]]$fit$distn, 
                           traits[[t]]$fit$parama,
                           traits[[t]]$fit$paramb),
           type = "mixed")

```
